---
title: "A Vignette to Bootstrap Panel Quantile Regression"
author: "Riki Saito"
date: "December 3, 2015"
output: html_document
---


First, we will view the summary of the example data used in this vignette.


```{r, echo=FALSE}
#Step 0 load data, packages, other useful functions
wd = "C:/Users/Riki/Dropbox/USGS/Work/bpqr/"
library(plm); library(quantreg)
data = read.csv(paste(wd, "bpqr_sample_data.csv",sep=""))
data = pdata.frame(data, index=c("Station","Year"))
row.prod <- function (x, na.rm = FALSE) {
  n <- nrow(x); y <- double(length = n)
  for (i in seq_len(n)) {
    y[i] <- prod(x[i,] , na.rm = na.rm)
  }
  y
}
summary(data)
```

Let us first examine the relationship of the response with the predictor, and the boxplot of the response by "firm" levels to motivate the use of a panel quantile regression.

```{r, fig.width = 7, fig.height = 7}
# Motivation for method
plot(data$Urban.Frac, log10(data$Original.Peaks), xlab="Urban.Frac",ylab="log10(Peaks)")
boxplot(log10(Original.Peaks) ~ Station, data)
```


We can argue the use of a quantile regression from the first plot, and the use of a panel regression from the second plot. Now let us try applying a panel quantile regression on this data. The first step is to perform a fixed effects panel regression on this data.

```{r, fig.width = 7, fig.height = 7}
formula = log10(Original.Peaks) ~ Urban.Frac;
plm  <- plm(formula,data=data,model="within");
prelim  <- cbind(coef(plm));    # This vector is K x 1 - only slopes.

hatalpha.i  <- cbind(fixef(plm));      	                  # dimension n x 1
all.firm <- gsub( "-.*$|\\..*$", "", row.names(data))
all.firm.fe <- merge(cbind(table(all.firm)),cbind(hatalpha.i), by="row.names")
intercept <- sum(row.prod(all.firm.fe[,-1]))/nrow(data)  #weighted mean of fixed effects
adj.hatalpha <- rep(hatalpha.i, table(all.firm)) - intercept
hist(hatalpha.i)
```

From the panel regression, we can see the varying intercepts of different "firms". We will remove this heterogeneity in the response, and the apply quantile regression.


```{r, fig.width = 7, fig.height = 7}
tau = c(.01,.05,seq(.1,.9,by=.1),.95,.99)
mf <- model.frame(formula, data)
y <- model.response(mf, "numeric")
x <- model.matrix(formula, data)
y.adj <- y - adj.hatalpha     #new y values adjusted for fixed effects
fm <- update.formula(formula, as.formula('y.adj ~ .'))
index = names(attr(data,"index"))
rqdf <- pdata.frame(cbind(data,y.adj),index=index)   #might need to edit
z <- rq(fm, data=rqdf, tau)
#z <- rq(y.adj~., data=data.frame(y.adj,x[,-1]), tau)
z$fe <- hatalpha.i
z

plot(data$Urban.Frac, y.adj, xlab="Urban.Frac")
plot(data$Urban.Frac, y.adj)
for(i in seq_len(length(tau))) abline(coef(z)[1,i],coef(z)[2,i])
```


Now with bootstrapping


```{r, echo=FALSE, fig.width = 7, fig.height = 7}
pqr <- function(formula, data, method="within", tau=.5){
  require(plm); require(quantreg);

  pmod  <- plm(formula,data=data,model=method)
  prelim  <- cbind(coef(pmod));    # This vector is K x 1 - only slopes.
  hatalpha.i  <- cbind(fixef(pmod));      	       # dimension n x 1, FE only
  all.firm <- gsub( "-.*$|\\..*$", "", row.names(data))
  all.firm.fe <- merge(cbind(table(all.firm)),cbind(hatalpha.i), by="row.names")
  intercept <- sum(row.prod(all.firm.fe[,-1]))/nrow(data)  #weighted mean of fixed effects
  adj.hatalpha <- rep(hatalpha.i, table(all.firm)) - intercept

  mf <- model.frame(formula, data)
  y <- model.response(mf, "numeric")
  x <- model.matrix(formula, data)
  y.adj <- y - adj.hatalpha     #new y values adjusted for fixed effects
  fm <- update.formula(formula, as.formula('y.adj ~ .'))
  index = names(attr(data,"index"))
  rqdf <- pdata.frame(cbind(data,y.adj),index=index)   #might need to edit
  z <- rq(fm, data=rqdf, tau)
  #z <- rq(y.adj~., data=data.frame(y.adj,x[,-1]), tau)
  z$fe <- hatalpha.i
  z
}

bpqr <- function(formula, data, B = 100, firm = 1, method="within", tau = .5){
  #cl <- match.call()
  #mf <- match.call(expand.dots = FALSE)

  require(plm); require(quantreg);
  pmod  <- plm(formula,data=data,model=method)
  prelim  <- cbind(coef(pmod));    # This vector is K x 1 - only slopes.
  hatalpha.i  <- cbind(fixef(pmod));      	       # dimension n x 1, FE only
  all.firm <- gsub( "-.*$|\\..*$", "", row.names(data))
  all.firm.fe <- merge(cbind(table(all.firm)),cbind(hatalpha.i), by="row.names")
  intercept <- sum(row.prod(all.firm.fe[,-1]))/nrow(data)  #weighted mean of fixed effects
  adj.hatalpha <- rep(hatalpha.i, table(all.firm)) - intercept

  mf <- model.frame(formula, data)
  y <- model.response(mf, "numeric")
  x <- model.matrix(formula, data)
  y.adj <- y - adj.hatalpha     #new y values adjusted for fixed effects
  fm <- update.formula(formula, as.formula('y.adj ~ .'))
  index = names(attr(data,"index"))
  rqdf <- pdata.frame(cbind(data,y.adj),index=index)   #might need to edit

  # Re-shape to wide to resample individual with their times
  key <- names(attr(rqdf,"index"))
  value <- all.vars(formula)
  dt <- pdata.frame(cbind(data[,c(key,value)],y.adj), index=key)

  firm.var <- key[firm]
  time.var <- key[-firm]

  widedt	<- reshape(dt, idvar=firm.var, timevar=time.var, direction="wide")

  n <- length(unique(data[,firm.var]))   # n panel samples
  K <- ncol(model.matrix(formula,data))  # parameters

  param <- array(0,dim=c(B,K,length(tau))) #initialize output data

  pb <- winProgressBar(title = "progress bar", min = 0, max = B, width = 300)
  Sys.time() -> start

  for (i in 1:B){
    # Generate random indices out of the n individuals
    samp <- sample(1:n,n,replace=T); # for balanced and unbalanced data

    wide_rdt <- widedt[samp,]    # bootstrap data
    wide_rdt[,firm.var] <- 1:n     # rename id to prevent duplicates

    # Reshape the bootstrap data back to long version to apply estimator
    rdt <- data.frame(na.omit(reshape(wide_rdt, direction="long")))
    names(rdt) <- names(dt)
    rdt <- rdt[order(rdt[firm.var], rdt[time.var]),]
    row.names(rdt) <- paste(rdt[,firm.var],rdt[,time.var],sep=".")
    rpdt    <- pdata.frame(rdt, index = key, row.names = T)

    # panel quant reg
    z <- pqr(formula, data=rpdt, tau=tau)

    # write result
    theta   <- coef(z)

    if(length(tau) == 1) param[i,,1] <- coef(z) else
    if(length(tau) > 1)  param[i,,] <- coef(z)

    Sys.sleep(0.1); setWinProgressBar(pb, i, title=paste( round(i/B*100, 0),"% done",sep=""))
  }
  close(pb)
  Sys.time() -> end
  cat("Bootstrap Complete. \nTime elapsed: ", round(difftime(end, start, units="secs"),3) ,"seconds")
  colnames(param) <- names(theta)
  return(param)
}

z2 = bpqr(formula, data, B = 10, firm = 1, tau = tau)
coef.z2 = apply(z2, 2:3, mean)

plot(data$Urban.Frac, y.adj, xlab="Urban.Frac")
for(i in seq_len(length(tau))) abline(coef(z)[1,i],coef(z)[2,i])
for(i in seq_len(length(tau))) abline(coef.z2[1,i],coef.z2[2,i],col="red")
```
